---
title: "Exploratory analysis of Amgen liver expression data"
output: html_document
---

### Summary

The goal of this analysis is simply to inspect the data, understand the overall
structure a bit better, and identify any interesting (or problematic) features
or samples in the data.

### Loading the data

The raw expression files (.CEL) were previously processed using the
`aroma.affymetrix` package. Specifically, all microarray samples were 
collectively processed with RMA for background correction, then quantile 
normalized. We can load the data again using `aroma` functions (this takes a
few moments, so we can turn on the `cache` option in the `r` chunk). 

**Note:** The `data` directory should be set as the working directory and
should include this `.Rmd` file for the following code to work.

```{r load_processed_data, cache=TRUE}
# Load the aroma.affymetrix library
library(aroma.affymetrix)

# Define inputs
chipType <- "Rat230_2"
projectName <- "amgenLiver"

# Define CDF
cdf <- AffymetrixCdfFile$byChipType(chipType)

# Load the original, raw CEL set
cs <- AffymetrixCelSet$byName(projectName, cdf = cdf)

# Define CEL set as RMA corrected, quantile normalized data
csN <- AffymetrixCelSet$byName(projectName, cdf = cdf, tags = "RBC,QN")
```

Next, we'll convert the `aroma` CEL set object into an `AffyBatch` object so
that we can use other functions in the `affy` package.

```{r convert_to_affybatch, cache=TRUE}
library(affy)

# Extract AffyBatch from CEL set object
liverData <- extractAffyBatch(csN)

# Also extract an AffyBatch object for the raw data
liverData_raw <- extractAffyBatch(cs)

```

Some basic plotting...

```{r simple_plots}
hist(liverData_raw)
boxplot(liverData_raw)
```

Use `affy` function `rma` to background correct and quantile normalize the raw
expression data (compare to `aroma` preprocessing steps).

```{r preprocess_data, cache=TRUE}
eset <- rma(liverData_raw) #11:19
hist(exprs(eset))
boxplot(exprs(eset))
```

Extract the expression data from the AffyBatch object...

```{r extract_eset}
eset <- exprs(liverData)
```

More basic plotting with the `eset` object...

```{r simple_plots_eset}
boxplot(eset)
```
